Idea in bullets

### 2022-06-10 paper/paper_in_bullets.tex
started acml paper

ALE plots are the best feature effect technique, because
blablabla. However, they have three drawbacks; Firstly, the do not
provide a measure of uncertainty about the effect they calculate (in
other words ). Secondly, they sometimes calculate different depending
on the number of bins (and we do not have a measure to choose which
one is closer to reality). Finally, for making accurate ALE plots, we
sometimes need to use variable-size bins.

@@

\begin{itemize}
\item \(K\), number of bins
\item \(dx\), bin-size
\item \(\texttt{min\_points\_per\_bin}\)
\end{itemize}
%
which have `1-1' relation:

\begin{equation*}
 dx \Leftrightarrow K \Leftrightarrow \texttt{min\_points\_per\_bin}
\end{equation*}
%
Therefore, we refer to them interchangeably.

@@

Altering the number of bins \( (K) \), leads to different ALE
plots. Depending on \(K\), we receive quite different ALE without any
indicator which one is the correct. In other words, we need a metric
to evaluate each plot.

@@

>>>>

ALE plots are the best feature effect (FE) technique, because ...
However, they have three drawbacks; Firstly, they do not provide
uncertainty about the measured effect, i.e.\ how confident to be that
the provided effect is the correct one. Secondly, in many cases, the
provide different effect depending on the number of bins (and we do
not have a measure to choose which one is closer to reality). Finally,
there are cases where it is impossible to create accurate ALE plots,
with fixed-size bins.

@@

\begin{itemize}
\item \(K\), number of bins
\item \(dx\), bin-size
\item \(\texttt{min\_points\_per\_bin}\)
\end{itemize}
%
There is `1-1' relation between these hyperparameters:

\begin{equation*}
 dx \leftrightarrow K \leftrightarrow \texttt{min\_points\_per\_bin}
\end{equation*}
%
Therefore, we refer to them interchangeably.

@@

Altering the number of bins \( (K) \) leads to different ALE plots and
we do not have any indicator which plot is closest to the truth.

@@

The combination of non-linear and linear parts makes variable-size
feature effect even more important.

\begin{itemize}
\item If we have enough samples, we need small bins to approach
  accurately the non-linear parts.
\item We need wider bins for robust estimation of the linear parts
\end{itemize}

\begin{equation} \label{eq:bullet-4-model}
  f(x_1, x_2) = 10x_1x_2 +
  \begin{cases}
    sinc(\pi x_1^2)  &,  0  \leq x_1 < \tau \\
    sinc(\pi \tau^2) &, \tau \leq x_1 \leq 5
  \end{cases}
\end{equation}
%
where \(x_1 \perp x_2 \Rightarrow p(x_1, x_2) = p(x_1)p(x_2)\) and \(\tau = 2.65\). The data generation distribution for \(x_1\) is:

\begin{equation} \label{eq:bullet-4-generative-x-1}
  p(x_1) =
  \begin{cases}
    \mathcal{U}(x_1; 0 , \tau)/2  &,  0  \leq x_1 < \tau \\
    \mathcal{N}(x_1; \mu = (5 + \tau)/2 , \sigma=0.3)/2 &, \tau \leq x_1 \leq 5
  \end{cases}
\end{equation}

and for \(x_2\):

\begin{equation} \label{eq:bullet-4-generative-x-2}
  p(x_2) = \mathcal{N}(x_2; \mu = 0 , \sigma=0.1)
\end{equation}

Therefore, the gradients wrt. \(x_1\) are:

\begin{equation} \label{eq:bullet-4-data-effect}
  \frac{\partial f}{ \partial x_1} (x_1) = 10x_2 +
  \begin{cases}
    2 \frac{\cos(\pi x_1^2)}{x_1} - 2 \frac{\sin(\pi x_1^2)}{\pi x_1^3}  &,  0  \leq x_1 < \tau \\
    0   &, \tau \leq x_1 \leq 5
  \end{cases}
\end{equation}
%
where \(x_2 \sim \mathcal{N}(\mu=0, \sigma=4)\). The ground truth ALE is:

\begin{equation} \label{eq:bullet-4-gt-ale}
  f_{\mathtt{ALE}}(x_s) = c +
  \begin{cases}
    sinc(\pi x_1^2)  &,  0  \leq x_1 < \tau \\
    sinc(\pi \tau^2)  &,  0  \leq x_1 < \tau \\
  \end{cases}
\end{equation}
%
where \(c\) is a normalizing constant. We generate N = 1000 data
points. In figure~\ref{fig:bullet-4-im-1}, we see the data points with
their respective gradients. In figure~\ref{fig:bullet-4-im-2}, we see
the fixed-size ALE plots; if we set \(K=50\), we capture the
non-linear parts in the first half, but we fail in linear second
part. Furthermore, we have empty bins, so we cannot trust the plot and
the standard error on the second half. The largest number of bins with
more than ten points in each bin, is \(K=5\) which fails to capture
the high-resolution effects. Variable size bins solves both problems
and provides the correct feature-effect.

\begin{figure}[!h]
  \centering
  \includegraphics[width=.49\linewidth]{bullet_4/im_1.png}
  \caption{Data points with their gradients}
  \label{fig:bullet-4-im-1}
\end{figure}

\begin{figure}[!h]
  \centering
  \includegraphics[width=.48\linewidth]{bullet_4/im_2.png}
  \includegraphics[width=.48\linewidth]{bullet_4/im_3.png}
  \caption{Fixed-size bins feature effect. Left: with K=50 and Right: with K=5.}
  \label{fig:bullet-4-im-2}
\end{figure}

\begin{figure}[!h]
  \centering
  \includegraphics[width=.75\linewidth]{bullet_4/im_4.png}
  \caption{Variable-size bins.}
  \label{fig:bullet-4-im-3}
\end{figure}
